Map{R}educe in {MPI} for Large-scale Graph Algorithms

### 2010-04-30 paper/graph.tex
Added discussion of improved SSSP algorithm
git-svn-id: svn+ssh://software-srn.sandia.gov/svn/mapreduce/trunk@597 3fc257bb-8550-0410-81b3-a5c103da83f7

As with PageRank and R-MAT generation, 
significant savings in communication and execution time
can be achieved by pre-aggregating certain MapReduce objects and 
storing data in more than one MapReduce object.  The improved algorithm
is shown in Figure~\ref{fig:sssp2}.

>>>>

As with PageRank and R-MAT generation, 
significant savings in communication and execution time
can be achieved by pre-aggregating certain MapReduce objects and 
storing data in more than one MapReduce object.  The improved algorithm
is shown in Figure~\ref{fig:sssp2}.  The MapReduce objects storing
vertices and edges are aggregated to processors only once at the beginning
of the computation.  Then, in each iteration of 
SSSP, only a small number of new candidate distances are generated.
The improved algorithm uses MapReduce object $MR$ to store these updates.
All communication is performed using $MR$, so the total amount of communication
needed is small compared to the original algorithm (which aggregates all
graph edges in each iteration).  Updated vertex distances are stored back
in the the MapReduce object for vertices.

\begin{figure}[htb]
 \begin{center}\fbox{\begin{minipage}{\textwidth} \begin{tabbing}
  xxxx\=xxxx\=xxxx\=xxxx\=xxxx\=xxxxxxxxxxxxxxxxxxxxxxxxxx\= \kill

Input:  \\
\> Source vertex $v_s$.\\
\> Pre-aggregated MapReduce object $V$ containing graph vertices $v_i$ with distances $d_i$ from $v_s$:  \\
\> \> Key = $v_i$; Value = $\infty$.\\
\> Pre-aggregated MapReduce object $E$ containing graph edges ($v_i, v_j$) with weights $w_{ij}$: \\
\> \>  Key = $v_i$; Value = $[v_j, w_{ij}]$ \\
$U$ = new MapReduce object to store updates to distances. \\
Map $U$:  Add $v_s$ with distance $d_s = 0$ to $U$. \\
\> Input:  $v_s$ \\
\> Output:  Key = $v_s$; Value = 0. \\
while not done: \\
\> done = 1; \\
\> Aggregate $U$. \\
\> Add $U$ to $V$. Empty $U$.\\
\> Convert $V$:  Key is vertex $v_i$. \\
\> Reduce $V$:  Pick best candidate distance for each vertex. \\
\> \> Input:  Key = $v_i$; Multivalue = candidate distances $d_i$. \\
\> \> Compute:  Find smallest candidate distance $d_{min}$ for $v_i$.  \\
\> \> \> If $d_{min} \ne d_i$, done = 0; \\
\> \> Output to $V$:  Key = $v_i$; Value = $d_{min}$. \\
\> \> Output to $U$: if $d_{min} \ne d_i$, Key = $v_i$; Value = $d_{min}$. \\
\> MPI\_Allreduce on done with MPI\_MIN. \\
\> if not done \\
\> \> Add $U$ to $E$.  Empty $U$. \\
\> \> Convert $E$. \\
\> \> Reduce $E$:  Generate candidate distances for neighbors of changed vertices. \\
\> \> \> Input:  Key = $v_i$; Multivalue = $[v_j, w_{ij}]$ for edges ($v_i, v_j$) \\
\> \> \> \> and distance $d_i$ if updated in previous step. \\
\> \> \> Output to $U$:  If updated distance $d_i$ exists, \\
\> \> \> \> Key = $v_j$; Value = $d_i+w_{ij}$ for each edge ($v_i, v_j$). \\
Output: MapReduce object $V$ with Key = $v_i$, Value = $d_i$.

 \label{fig:sssp2}
\end{figure}
